EDA of Wages Tidied Data

### 2020-12-22 EDA_wages_hs2020.Rmd
Add GAM model

library(tidyverse)
library(brolgar)
library(visdat)
library(naniar)
library(skimr)
library(kableExtra)
library(patchwork)
library(gghighlight)
library(stickylabeller)
```

@@

ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_strata(n_strata = 100)

@@

```{r}
ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  facet_strata(along = -year)
```
Learn: Most of the samples participated in the study in 1979, this is also inline with the finding in Figure \@ref(fig:first-time). We can see that generally the subjects of the study have the same starting wages but it became vary in the end of the study.

>>>>

library(tidyverse)
library(brolgar)
library(visdat)
library(naniar)
library(skimr)
library(kableExtra)
library(patchwork)
library(gghighlight)
library(stickylabeller)
library(mgcv)
library(modelr)
library(tsibble)
```

@@

set.seed(2020)
ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_strata(n_strata = 100)

@@

```{r}
set.seed(2020)
ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  facet_strata(along = -year)
```
Learn: Most of the samples participated in the study in 1979, this is also inline with the finding in Figure \@ref(fig:first-time). We can see that generally the subjects of the study have the same starting wages but it became vary in the end of the study.

## Build GAM model

```{r id-fct}
wages <- wages %>%
  mutate(id_fct = as.factor(id),
         year0 = year - 1979)
```

```{r eval = FALSE}
# wages_gam <- gam(
#   mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
#   data = wages,
#   method = "REML"
# )
```

Since we have a quite large data set, we have to subset it first to model the data. Here, we would like to build a Generalized Additive Model (GAM) model, by randomly subset the daya the data using brolgar's` `sample_frac_keys` function.

```{r}
set.seed(2020)
wages_sample <- sample_frac_keys(wages, size = 0.01)

```{r}
wages_gam <- gam(
  mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
  data = wages_sample,
  method = "REML"
)
```

```{r}
wages_aug <- wages_sample %>%
  add_predictions(wages_gam, var = "pred") %>%
  add_residuals(wages_gam, var = "res") %>%
  group_by_key() %>%
  mutate(rss = sum(res^2)) %>%
  ungroup()
```

```{r}
set.seed(2020)
wages_aug %>%
  sample_n_keys(12) %>%
  ggplot(aes(x = year,
             y = pred,
             group = id)) +
  geom_line(color = "steelblue") +
  geom_point(aes(y = mean_hourly_wage)) +
  facet_wrap(~id)
```
